Game-theoretic strengthenings of Menger's property

### 2014-07-15 phd/talks-and-presentations/2014-07-23-menger-talk.tex
draft of talk on menger game for summer conference

@@

% Strategy uparrow shortcuts
\newcommand{\win}{\uparrow}
\newcommand{\prewin}{\underset{\text{pre}}{\uparrow}}
\newcommand{\markwin}{\underset{\text{mark}}{\uparrow}}
\newcommand{\tactwin}{\underset{\text{tact}}{\uparrow}}
\newcommand{\kmarkwin}[1]{\underset{#1\text{-mark}}{\uparrow}}
\newcommand{\ktactwin}[1]{\underset{#1\text{-tact}}{\uparrow}}
\newcommand{\codewin}{\underset{\text{code}}{\uparrow}}
\newcommand{\limitwin}{\underset{\text{limit}}{\uparrow}}

@@

\newcommand{\al}[1]{{#1}^*}

@@

\newcommand{\alcomp}{\al\parallel}

@@

  Sketch of Scheeper's proof:
  \begin{itemize}
    \item Using second-countability and the winning strategy for $\pl F$,
          construct certain subsets $R_s$ for $s \in \omega^{<\omega}$ such
          that $X = \bigcup_{s\in\omega^{<\omega}}$.
    \item Using regularity, show that each $R_s$ is compact.
    \item The result follows since $|\omega^{<\omega}|=\omega$.
  \end{itemize}

@@

\begin{frame}
  foo
\end{frame}

@@

\begin{frame}
Questions? Thanks for having me!
\end{frame}

>>>>

\usepackage{tikz}
\usetikzlibrary{matrix} % for diagrams

@@

% Strategy uparrow shortcuts
\newcommand{\win}{\uparrow}
\newcommand{\prewin}{\underset{\text{pre}}{\uparrow}}
\newcommand{\markwin}{\underset{\text{mark}}{\uparrow}}
\newcommand{\tactwin}{\underset{\text{tact}}{\uparrow}}
\newcommand{\kmarkwin}[1]{\underset{#1\text{-mark}}{\uparrow}}
\newcommand{\ktactwin}[1]{\underset{#1\text{-tact}}{\uparrow}}
\newcommand{\codewin}{\underset{\text{code}}{\uparrow}}
\newcommand{\limitwin}{\underset{\text{limit}}{\uparrow}}
\newcommand{\notprewin}{\underset{\text{pre}}{\not\uparrow}}
\newcommand{\notmarkwin}{\underset{\text{mark}}{\not\uparrow}}
\newcommand{\nottactwin}{\underset{\text{tact}}{\not\uparrow}}
\newcommand{\notkmarkwin}[1]{\underset{#1\text{-mark}}{\not\uparrow}}
\newcommand{\notktactwin}[1]{\underset{#1\text{-tact}}{\not\uparrow}}
\newcommand{\notcodewin}{\underset{\text{code}}{\not\uparrow}}
\newcommand{\notlimitwin}{\underset{\text{limit}}{\not\uparrow}}

@@

@@

\newcommand{\al}[1]{{#1}^*}
\newcommand{\alcomp}{\wr}

@@

  Sketch of Scheeper's proof:
  \begin{itemize}
    \item Using second-countability and the winning strategy for $\pl F$,
          construct certain subsets $K_s$ for $s \in \omega^{<\omega}$ such
          that $X = \bigcup_{s\in\omega^{<\omega}}K_s$.
    \item Using regularity, show that each $K_s$ is compact.
    \item The result follows since $|\omega^{<\omega}|=\omega$.
  \end{itemize}

@@

\begin{frame}
  Knowledge of two previous moves versus one is an important distinction:
  in the former case, the player is able to react to change by the
  opponent.

  \begin{definition}
    Let $\oneptlind\kappa = \kappa\cup\{\infty\}$ be the
    \term{one point Lindel\"of-ication} of discrete $\kappa$: neighborhoods
    of $\infty$ are exactly the co-countable sets containing it.
  \end{definition}

  $\oneptlind\kappa$ is a simple space which is a regular and Lindel\"of, but
  not second-countable space or $\sigma$-compact. Thus
  $\pl F\notkmarkwin{1}\mengame{\oneptlind\kappa}$, but it's easy to see
  that $\pl F\win\mengame{\oneptlind\kappa}$. What about $2$-Mark\"ov
  strategies?
\end{frame}

\begin{frame}
  In 1991, Scheepers introduced the statement $\alcompS\kappa$ to study an
  infinite game involving the countable and finite subsets of $\kappa$.

  \begin{game}
    Let $\fillgameS\kappa$ denote the \term{strict union filling game}
    with two players $\pl C$, $\pl F$. In round $0$, $\pl C$ chooses
    $C_0\in[\kappa]^{\leq\omega}$, followed by $\pl F$ choosing
    $F_0\in[\kappa]^{<\omega}$. In round $n+1$, $\pl C$ chooses
    $C_{n+1}\in[\kappa]^{\leq\omega}$ such that $C_{n+1}\supset C_n$, followed
    by $\pl F$ choosing $F_{n+1}\in[\kappa]^{<\omega}$.

    $\pl F$ wins the game if
    $\bigcup_{n<\omega} F_n\supseteq\bigcup_{n<\omega} C_n$; otherwise, $\pl C$
    wins.
  \end{game}
\end{frame}

\begin{frame}
  \begin{definition}
    For two functions $f,g$ we say $f$ is \term{almost compatible} with
    $g$ ($f\alcomp g$) if $|\{x\in\dom(f)\cap\dom(g):f(x)\not=g(x)\}|<\omega$.
  \end{definition}
  \pause
  \begin{definition}
    $\alcompS\kappa$ states that there exist functions
    $f_A:A\to\omega$ for each $A\in[\kappa]^{\leq\omega}$ such that
    $|f_A^{-1}(n)|<\omega$ for all $n<\omega$ and
    $f_A\alcomp f_B$ for all $A,B\in[\kappa]^\omega$.
  \end{definition}
  \pause
  \begin{theorem}
    $\alcompS{\omega_1}$; $Con(\alcompS{2^\omega}+\neg CH)$;
    $\neg\alcompS\kappa$ for $\kappa>2^\omega$.
  \end{theorem}
\end{frame}

\begin{frame}
  The round number is unnecessary in Scheeper's game, since $\pl C$ must
  choose strictly increasing sets.

  \begin{theorem}
    If $\alcompS\kappa$, then $\pl F\ktactwin2 \fillgameS\kappa$.
  \end{theorem}

  As it turns out, a related game characterizes $\mengame{\oneptlind\kappa}$.

  \begin{definition}
    Let $\fillgameInt\kappa$ denote the \term{intersection filling game}
    analogous to $\fillgameS\kappa$, except that $\pl C$ has no restriction
    on the countable sets she choooses, but $\pl F$ need only ensure that
    $\bigcup_{n<\omega}\mc F_n\supseteq \bigcap_{n<\omega}\mc U_n$ to win
    the game.
  \end{definition}
\end{frame}

\begin{frame}
  \begin{theorem}
    $\pl F \kmarkwin{k} \mengame{\oneptlind\kappa}$ if and only if
    $\pl F \kmarkwin{k} \fillgameInt\kappa$.
  \end{theorem}

  \begin{theorem}
    If $\alcompS\kappa$, then $\pl F\kmarkwin{2} \fillgameInt\kappa$.
  \end{theorem}

  \begin{proof}
    Let $f_A:A\to\omega$ witness $\alcompS\kappa$. Then we define the
    winning $2$-Mark\"ov strategy $\sigma$ as follows:
      \[
        \sigma(\<A\>,0) = \{\alpha\in A: f_A(\alpha) = 0\}
      \]
      \[
        \sigma(\<A,B\>,n+1)
          =
        \{\alpha\in A\cap B
          :
        f_B(\alpha) \leq n+1
          \text{ or }
        f_A(\alpha)\not=f_B(\alpha)\}
      \]
  \end{proof}
\end{frame}

\begin{frame}
  \begin{corollary}
    $\pl F\kmarkwin{2}\mengame{\oneptlind\omega_1}$, but
    $\pl F\notkmarkwin1\mengame{\oneptlind\omega_1}$.
  \end{corollary}
  \begin{figure}
    \includegraphics[height=1.7in]{mengerGameChart.png}
  \end{figure}
\end{frame}

\section{Quesitons}

\begin{frame}
  \begin{question}
    Does $\pl F\kmarkwin2\mengame{\oneptlind\kappa}$ imply $\alcompS\kappa$?
  \end{question}

  \begin{question}
    Are $\pl F\win\mengame X$ and $\pl F\kmarkwin2\mengame X$ distinct?
  \end{question}

  An affirmative answer to the first question answers this since
  $\neg\alcompS\kappa$ for $\kappa>2^\omega$.
\end{frame}
\begin{frame}
  \begin{question}
    Where does $\pl F\kmarkwin2\mengame X$ fit in with other properties
    between $\sigma$-(relatively-)compact and Menger?
  \end{question}
  $\pl F\kmarkwin2\mengame X$ seems to characterize an
  ``almost-$\sigma$-(relative-)compactness''.

  Any sufficient property must imply $\pl F\win\mengame X$ (ruling out
  Hurewicz[?]), and any (interesting) necessary property shouldn't be implied by
  $\pl F\win\mengame X$ (ruling out Alster and productively Lindel\"of for
  regular spaces).
\end{frame}

@@

\begin{frame}
Questions? Thanks for listening!
\end{frame}
